---
title: "244 Lab 7"
author: "Gage Clawson"
date: "2/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(tmap)
library(sf)
library(spatstat)
library(maptools)
library(sp)
library(raster)
library(gstat)
```

Part 1. Hawaii raster practice
```{r}
hi_par <- raster("PAR_CLIM_M.tif")

hi_sst <- raster("SST_LTM.tif")

hi_chl <- raster("CHL_LTM.tif")

par(mfrow = c(1,3))
plot(hi_par)
plot(hi_sst)
plot(hi_chl)

```

Reproject to WGS84:
```{r}

wgs84 <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +no_defs"

hi_sst_84 <- projectRaster(hi_sst, crs = wgs84, method = "bilinear")

hi_sst_84@crs

plot(hi_sst_84)
```

raster::aggregate() for resampling
```{r}
#resolution decreased by an order of 10. Think hard before doing this. 
sst_rs <- aggregate(hi_sst, fact = 10)
plot(sst_rs)

```

Crop a raster:
```{r}
hi_sst_84@extent

#create a bounding box. Decided on this box because it isolates one island. 
bounds <- as(extent(-156.2, -154.5, 18.7, 20.4), 'SpatialPolygons')

#if something does not have a crs, this is the way to do it. otherwise have to use projectRaster.
#here we set bounds crs to the same of hi_sst_84
crs(bounds) <- crs(hi_sst_84)

#crop sst info
sst_crop <- crop(hi_sst_84, bounds)
plot(sst_crop)
```

Simple raster math:

Nonsensical variable called "tropicality" that is the sum of PAT + SST + 2*ChlA, and we want to map that variable...
```{r}
hi_par_84 <- projectRaster(hi_par, crs = wgs84, method = "bilinear")

hi_chla_84 <- projectRaster(hi_chl, crs = wgs84, method = "bilinear")

par(mfrow = c(1,3))
plot(hi_par_84)
plot(hi_sst_84)
plot(hi_chla_84)


```

```{r}
trop <- hi_par_84 + hi_sst_84 + 2*hi_chla_84

plot(trop)
```

Now let's try to look at something using tmap:
```{r}
islands <- read_sf(dsn = 'islands', layer = "Island_boundaries") %>%
  dplyr::select(Island) %>% 
  st_simplify(dTolerance = 10) %>% 
  st_transform(crs = 4326)
```

```{r}
tmap_mode("plot")

sst_map <- tm_shape(hi_sst_84) +
  tm_raster(title = "Mean Sea Surface Temperature") +
  tm_layout(bg.color = "navyblue", legend.position = c("left", "bottom"), legend.text.color ="white", legend.text.size = 0.5) + 
  tm_shape(islands) +
  tm_fill("darkgreen")


tmap_save(sst_map, "gage_hawaii.png", height = 5)
```

Condition rasters adnd masking

Let's say: we have a sensitive species and we're trying to find habitat that it might like. We know: they like warm water (average SST >= 25.6 degrees) and solar (PAR) below 54. 

```{r}
extent(hi_sst_84) <- extent(hi_par_84)

cr <- raster(nrow = 822, ncol = 1229, xmn = -160.4365, xmx = -154.5373, ymn = 18.7309, ymx = 22.44634)


sst_new <- resample(hi_sst_84, cr, method = "bilinear")

compareRaster(sst_new, hi_par_84)

plot(sst_new)
plot(hi_par_84)

#make a cropped version just for Kauai
bounds_main <- as(extent(-159.9, -159.2, 21.7, 22.3), 'SpatialPolygons')
crs(bounds_main) <- crs(sst_new)

par_kauai <- crop(hi_par_84, bounds_main)
sst_kauai <- crop(sst_new, bounds_main)

plot(sst_kauai)
```

Now we only want to isolate regions where temp >= 25.4, PAR < 54
```{r}
par_hab <- par_kauai
par_hab[par_hab >= 54] <- NA

sst_hab <- sst_kauai
sst_hab[sst_hab < 25.4] <- NA

plot(par_hab)

# Where do these overlap? raster::mask
suit_hab <- mask(sst_hab, par_hab)

plot(suit_hab)
```


###Part 2. Point pattern analysis

An analysis of red tree voles in Humboldt County

```{r}
voles <- read_sf(dsn = "redtreevoledata", layer = "ds033") %>%
  dplyr::select(COUNTY) %>%
  filter(COUNTY == "HUM") %>%
  st_transform(crs = 4326)

plot(voles)

#Get humboldt county outline

humboldt <- read_sf(dsn = 'redtreevoledata' , layer = "california_county_shape_file") %>%
  filter(NAME == "Humboldt") %>%
  dplyr::select(NAME)

st_crs(humboldt) <- 4326

tm_shape(humboldt) +
  tm_fill() + 
  tm_shape(voles) +
  tm_dots(size = 0.2)

ggplot() +
  geom_sf(data = humboldt) +
  geom_sf(data = voles)

ggsave("humvoles.png", units = "in" , width = 4, height = 6, dpi = 300)

```


We want to explore point patterns a few different ways:

- wuadrat analysis
- distance based (neighbor analysis, G-function and K-function)

```{r}
voles_sp <- as(voles, 'Spatial')
voles_ppp <- as(voles_sp, "ppp")

humboldt_sp <- as(humboldt, "Spatial") 
humboldt_win <- as(humboldt_sp, "owin")

voles_pb <- ppp(voles_ppp$x, voles_ppp$y, window = humboldt_win)

plot(voles_pb)

```


Quadrat test:
```{r}
voles_qt <- quadrat.test(voles_pb, nx = 5, ny =10)

voles_qt

plot(voles_pb)
#plot(voles_qt, add = TRUE, cex = 0.4)
```


plotting kernel densities for spatial data 

```{r}
#sigma changes the bandwidth. always report the bandwidth. 
point_density <- density(voles_pb, sigma = 0.02)
plot(point_density)

vole_raster <- raster(point_density, crs = wgs84)

tm_shape(vole_raster) +
  tm_raster()

```


Nearest neighbor

G-function: considers the distance of each observation to its NEAREST neighbor

K-function: considers how close all neighboring observations are to an event (concentric circles)

```{r}
r <- seq(0, 0.15, by = 0.005)

gfunction <- envelope(voles_pb, fun = Gest, r = r, nsim = 20)

plot(gfunction)

plot(gfunction$obs ~ gfunction$r, type = "l", col = "black") +
lines(gfunction$theo ~ gfunction$r, type = "l", col = "red")


#K/L function: asking about how close ALL neighbors are to every event in the spatial window

r2 <- seq(0, 0.5, by = 0.05)

lfunction <- envelope(voles_pb, fun = Lest, r = r2, nsim = 20, global = TRUE)

plot(lfunction$obs ~ lfunction$r, type = "l", col = "blue") +
lines(lfunction$theo ~ lfunction$r, type = "l", col = "red")

```

Diggle-Cressie-Loosmore-Ford test of CSR

```{r}
DCLFTest <- dclf.test(voles_pb, nsim = 30)
DCLFTest

```


###Part 3. Spatial interpolation by kriging

```{r}
ks_rain <- read_csv("KSRain2.csv")

ks_sf <- st_as_sf(ks_rain, coords = c("LON", "LAT"), crs = 4326)
plot(ks_sf)

ks_counties <- read_sf(dsn = 'KSCounties', layer = "ks_counties_shapefile")

st_crs(ks_counties) <- 4326
plot(ks_counties)

tm_shape(ks_counties) + 
  tm_fill() +
  tm_shape(ks_sf) +
  tm_dots("AMT", size = 0.5)

```

```{r}
ks_sp <- as_Spatial(ks_sf)


```

Make a spatial grid to interpolate values over
```{r}
#she chose some of these for resolution. For sake of time don't go much higher 

lat <- seq(37, 40, length.out = 200)
long <- seq(-94.6, -102, length = 200)

grid <- expand.grid(lon = long, lat = lat)
#convert into an af object with coords
grid_sf <- st_as_sf(grid, coords = c("lon", "lat"), crs = 4326)

grid_sp <- as_Spatial(grid_sf)

```

Then make a variogram and find the variogram model
```{r}
ks_vgm <- variogram(AMT ~ 1, ks_sp)

plot(ks_vgm)

ks_vgm_fit <- fit.variogram(ks_vgm, model = vgm(nugget = 0.2, psill = 0.8, range = 200, model = "Sph"))

plot(ks_vgm, ks_vgm_fit)
```

Now we can do spatial interpolation

```{r}


ks_krige <- krige(AMT ~ 1, ks_sp, grid_sp, model = ks_vgm_fit)


ks_krige_df <- as.data.frame(ks_krige)

ks_krige_2 <- ks_krige_df %>%
  rename(lon = coords.x1, lat = coords.x2, predicted = var1.pred, err = var1.var)

rain_predicted <- st_as_sf(ks_krige_2, coords = c("lon", "lat"), crs = 4326)

#get outline of kansas and crop to that. It is in folder called states. 
ks <- read_sf(dsn = "states", layer = "cb_2017_us_state_20m") %>%
  dplyr::select(NAME) %>% 
  filter(NAME == "Kansas") %>%
  st_transform(crs = 4326)

plot(rain_predicted)

rain_cropped <- st_intersection(rain_predicted, ks)

plot(rain_cropped)
```
